---
title: "Hierarchical PG Analyses: Primary, Sensitivity, and Severity Models"
author: "OpenAI"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Overview

This notebook revises the original `code/analysis/logit.Rmd` workflow to better match the current analytic plan.

### Primary decisions

1. Use a **hierarchical logistic regression** to predict **PG vs HC**.
2. Keep **moderators in Block 4** only.
3. Treat **DOSPERT as domain-specific** in the primary model.
4. **Exclude the gambling domain** from the primary model because of likely criterion overlap with PG status.
5. Treat **investment/financial risk** cautiously. By default, it is excluded from the primary model, but a switch is included below in case you want to add it.
6. Run two follow-up models:
   - a **non-gambling DOSPERT composite** sensitivity model,
   - an **exploratory gambling-domain** model.
7. Run a **within-PG severity analysis** predicting GSAS total.

## Packages

```{r packages}
library(dplyr)
library(tidyr)
library(ggplot2)
library(car)
library(broom)
library(pscl)
library(lmtest)
library(sandwich)
library(knitr)
library(purrr)
```

## Project paths and data import

```{r paths}
# Helper: find the repo root whether knitting from the repo root or from code/analysis/
find_project_root <- function() {
  candidates <- unique(c(
    normalizePath(getwd(), winslash = "/", mustWork = FALSE),
    normalizePath(file.path(getwd(), ".."), winslash = "/", mustWork = FALSE),
    normalizePath(file.path(getwd(), "..", ".."), winslash = "/", mustWork = FALSE)
  ))

  has_required_files <- function(path) {
    file.exists(file.path(path, "derivatives", "scoring", "scoring-mastersheet.csv")) &&
      file.exists(file.path(path, "derivatives", "scoring", "geo-variables.csv"))
  }

  hit <- candidates[vapply(candidates, has_required_files, logical(1))]
  if (length(hit) == 0) {
    stop("Could not locate the project root. Knit this file from the repo root or code/analysis/.")
  }
  hit[[1]]
}

project_root <- find_project_root()
score_path   <- file.path(project_root, "derivatives", "scoring", "scoring-mastersheet.csv")
geo_path     <- file.path(project_root, "derivatives", "scoring", "geo-variables.csv")

df     <- read.csv(score_path)
df_geo <- read.csv(geo_path)
```

## Analysis switches

```{r switches}
# Set to TRUE if you want to keep DOSPERT investment/financial risk in the primary model.
include_investment_primary <- FALSE

# Keep this TRUE to match the original repo's outlier handling for relative sensitivity.
apply_original_reward_outlier_filter <- TRUE
```

## Helper functions

```{r helpers}
z_score <- function(x) as.numeric(scale(x))

make_or_table <- function(model) {
  est <- coef(model)
  ci  <- suppressMessages(confint(model))
  data.frame(
    term = names(est),
    OR = exp(est),
    CI_low = exp(ci[, 1]),
    CI_high = exp(ci[, 2]),
    row.names = NULL
  )
}

make_formula <- function(outcome, block_terms, interactions = NULL) {
  rhs <- c(unlist(block_terms), interactions)
  stats::as.formula(paste(outcome, "~", paste(rhs, collapse = " + ")))
}

fit_block_models <- function(data, outcome, blocks, interactions = NULL, family = binomial()) {
  formulas <- list(
    m1 = make_formula(outcome, blocks[1]),
    m2 = make_formula(outcome, blocks[1:2]),
    m3 = make_formula(outcome, blocks[1:3]),
    m4 = make_formula(outcome, blocks[1:3], interactions = interactions)
  )

  models <- purrr::map(formulas, ~ glm(.x, data = data, family = family, control = glm.control(maxit = 100)))
  list(models = models, formulas = formulas)
}

compare_block_models <- function(model_list) {
  list(
    block_1_vs_2 = anova(model_list$m1, model_list$m2, test = "Chisq"),
    block_2_vs_3 = anova(model_list$m2, model_list$m3, test = "Chisq"),
    block_3_vs_4 = anova(model_list$m3, model_list$m4, test = "Chisq"),
    AIC = AIC(model_list$m1, model_list$m2, model_list$m3, model_list$m4),
    BIC = BIC(model_list$m1, model_list$m2, model_list$m3, model_list$m4),
    pR2 = tibble(
      model = c("m1", "m2", "m3", "m4"),
      McFadden = c(
        pR2(model_list$m1)["McFadden"],
        pR2(model_list$m2)["McFadden"],
        pR2(model_list$m3)["McFadden"],
        pR2(model_list$m4)["McFadden"]
      )
    )
  )
}
```

## Preprocessing

```{r preprocess}
# Reward sensitivity composites from the original repo
# (kept here so this notebook stays aligned with Cooper's prior workflow)
df$bisbas_bas <- df$bisbas_drive + df$bisbas_fun_seek + df$bisbas_reward_res

df$bisbas_bas_z_score <- z_score(df$bisbas_bas)
df$spsrq_sr_z_score   <- z_score(df$spsrq_sr)
df$rs_standard        <- df$bisbas_bas_z_score - df$spsrq_sr_z_score

if (apply_original_reward_outlier_filter) {
  q3 <- quantile(df$rs_standard, 0.75, na.rm = TRUE)
  q1 <- quantile(df$rs_standard, 0.25, na.rm = TRUE)
  iqr_value <- q3 - q1
  upper_bound <- q3 + 1.5 * iqr_value
  lower_bound <- q1 - 1.5 * iqr_value

  df <- df[!(df$rs_standard >= upper_bound | df$rs_standard <= lower_bound) | is.na(df$rs_standard), ]
}

df$rew_composite <- df$spsrq_sr_z_score + df$bisbas_bas_z_score

# Merge geographic/contextual variables
# Note: use the geo file already in the repo rather than refetching ACS data.
df <- df %>%
  left_join(df_geo, by = "ResponseId") %>%
  mutate(across(everything(), ~ ifelse(. == -999, NA, .)))

# Define gambling groups cleanly
# HC: non-gambling / minimal symptoms
# PG: problem gambling criteria met
# Mid: everyone in between, excluded from binary case-control analyses
df <- df %>%
  mutate(
    pg_group = case_when(
      sogs_total >= 1 & bbgs_total >= 4 & (mandf_casinog + mandf_onlineg) >= 1 ~ "PG",
      sogs_total == 0 & bbgs_total <= 3 & (mandf_casinog + mandf_onlineg) == 0 ~ "HC",
      TRUE ~ "Mid"
    ),
    pg_group = factor(pg_group, levels = c("HC", "Mid", "PG")),
    PG = ifelse(pg_group == "PG", 1L, 0L),
    gender = factor(demo_gender, levels = c(1, 2), labels = c("Male", "Female")),
    income_median_split = factor(
      ifelse(ses_tpi <= 2, "Below Median (<$40k)", "Above Median (≥$40k)"),
      levels = c("Below Median (<$40k)", "Above Median (≥$40k)")
    )
  )
```

## Build the HC vs PG analytic dataset

```{r build-binary-dataset}
# Start from the case-control comparison only
# and require complete data on all variables needed downstream.
dat_binary <- df %>%
  filter(pg_group %in% c("HC", "PG")) %>%
  drop_na(
    PG, gender, income_median_split,
    TotalCrime, unemployed, Gini,
    demo_yrs,
    rew_composite,
    dospert_gam, dospert_inv, dospert_hs, dospert_soc, dospert_ethic, dospert_rec,
    susd_depression, susd_mania, scared_total, uclal_total,
    gsas_total
  ) %>%
  mutate(
    income = income_median_split,

    # Context
    z_crime = z_score(TotalCrime),
    z_unemp = z_score(unemployed),
    z_gini  = z_score(Gini),
    z_neigh_disadv = rowMeans(cbind(z_crime, z_unemp, z_gini), na.rm = TRUE),

    # Demographics
    z_age = z_score(demo_yrs),

    # Reward sensitivity
    z_reward = z_score(rew_composite),

    # DOSPERT domains
    z_dospert_gam   = z_score(dospert_gam),
    z_dospert_inv   = z_score(dospert_inv),
    z_dospert_hs    = z_score(dospert_hs),
    z_dospert_soc   = z_score(dospert_soc),
    z_dospert_ethic = z_score(dospert_ethic),
    z_dospert_rec   = z_score(dospert_rec),

    # Clinical / social
    z_depression = z_score(susd_depression),
    z_mania      = z_score(susd_mania),
    z_anxiety    = z_score(scared_total),
    z_loneliness = z_score(uclal_total),

    # Sensitivity variants
    dospert_nongam_primary_raw = if (include_investment_primary) {
      rowMeans(cbind(dospert_hs, dospert_soc, dospert_ethic, dospert_rec, dospert_inv), na.rm = FALSE)
    } else {
      rowMeans(cbind(dospert_hs, dospert_soc, dospert_ethic, dospert_rec), na.rm = FALSE)
    },
    z_dospert_nongam_comp = z_score(dospert_nongam_primary_raw),
    z_dospert_total       = z_score(dospert_gam + dospert_inv + dospert_hs + dospert_soc + dospert_ethic + dospert_rec)
  ) %>%
  filter(gender %in% c("Male", "Female"))

cat("Binary analytic N =", nrow(dat_binary), "\n")
cat("HC =", sum(dat_binary$PG == 0), "\n")
cat("PG =", sum(dat_binary$PG == 1), "\n")
```

## Inspect DOSPERT overlap before modeling

```{r dospert-overlap}
dospert_terms_for_corr <- c(
  "z_dospert_hs", "z_dospert_soc", "z_dospert_ethic", "z_dospert_rec", "z_dospert_inv", "z_dospert_gam"
)

round(cor(dat_binary[, dospert_terms_for_corr], use = "pairwise.complete.obs"), 2)
```

## Define model blocks

```{r define-blocks}
primary_dospert_terms <- c("z_dospert_hs", "z_dospert_soc", "z_dospert_ethic", "z_dospert_rec")
if (include_investment_primary) {
  primary_dospert_terms <- c(primary_dospert_terms, "z_dospert_inv")
}

block_1 <- c("z_neigh_disadv", "gender", "z_age", "income")
block_2_primary <- c("z_reward", primary_dospert_terms)
block_2_comp    <- c("z_reward", "z_dospert_nongam_comp")
block_2_total   <- c("z_reward", "z_dospert_total")
block_2_explor  <- c("z_reward", primary_dospert_terms, "z_dospert_gam")
block_3 <- c("z_depression", "z_mania", "z_anxiety", "z_loneliness")
block_4_interactions <- c("z_reward:z_loneliness", "z_reward:z_age")
```

## Primary hierarchical logistic regression: non-gambling DOSPERT domains

```{r fit-primary-logit}
primary_fit <- fit_block_models(
  data = dat_binary,
  outcome = "PG",
  blocks = list(block_1, block_2_primary, block_3),
  interactions = block_4_interactions,
  family = binomial(link = "logit")
)

m1_primary <- primary_fit$models$m1
m2_primary <- primary_fit$models$m2
m3_primary <- primary_fit$models$m3
m4_primary <- primary_fit$models$m4
```

```{r compare-primary-logit}
compare_primary <- compare_block_models(primary_fit$models)
compare_primary$block_1_vs_2
compare_primary$block_2_vs_3
compare_primary$block_3_vs_4
compare_primary$AIC
compare_primary$BIC
compare_primary$pR2
```

```{r primary-vif}
# Main-effects VIF is the most interpretable first pass.
vif(m3_primary)

# For the interaction model, use predictor-level VIF.
vif(m4_primary, type = "predictor")
```

```{r primary-or-table}
primary_or <- make_or_table(m4_primary)
primary_or
```

```{r primary-robust-se}
coeftest(m4_primary, vcov = vcovHC(m4_primary, type = "HC3"))
```

```{r primary-clean-table}
tidy(m4_primary, exponentiate = TRUE, conf.int = TRUE) %>%
  mutate(
    across(c(estimate, conf.low, conf.high), ~ round(., 3)),
    p.value = round(p.value, 3),
    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      p.value < .10  ~ ".",
      TRUE ~ ""
    )
  ) %>%
  rename(OR = estimate, CI_low = conf.low, CI_high = conf.high) %>%
  kable()
```

## Sensitivity model: non-gambling DOSPERT composite

```{r fit-composite-logit}
comp_fit <- fit_block_models(
  data = dat_binary,
  outcome = "PG",
  blocks = list(block_1, block_2_comp, block_3),
  interactions = block_4_interactions,
  family = binomial(link = "logit")
)

m4_comp <- comp_fit$models$m4
compare_comp <- compare_block_models(comp_fit$models)
compare_comp$block_1_vs_2
compare_comp$block_2_vs_3
compare_comp$block_3_vs_4
compare_comp$AIC
compare_comp$pR2
```

## Exploratory model: add the gambling domain back in

```{r fit-exploratory-gambling-domain}
explor_fit <- fit_block_models(
  data = dat_binary,
  outcome = "PG",
  blocks = list(block_1, block_2_explor, block_3),
  interactions = block_4_interactions,
  family = binomial(link = "logit")
)

m4_explor <- explor_fit$models$m4
compare_explor <- compare_block_models(explor_fit$models)
compare_explor$block_1_vs_2
compare_explor$block_2_vs_3
compare_explor$block_3_vs_4
compare_explor$AIC
compare_explor$pR2
```

## Optional comparison: total DOSPERT instead of domains

```{r fit-total-dospert-model}
total_fit <- fit_block_models(
  data = dat_binary,
  outcome = "PG",
  blocks = list(block_1, block_2_total, block_3),
  interactions = block_4_interactions,
  family = binomial(link = "logit")
)

m4_total <- total_fit$models$m4
compare_total <- compare_block_models(total_fit$models)
compare_total$AIC
compare_total$pR2
```

## Compare final models side by side

```{r final-model-comparison}
model_compare <- tibble(
  model = c(
    "Primary: non-gambling domains",
    "Sensitivity: non-gambling composite",
    "Exploratory: add gambling domain",
    "Optional: total DOSPERT"
  ),
  AIC = c(
    AIC(m4_primary),
    AIC(m4_comp),
    AIC(m4_explor),
    AIC(m4_total)
  ),
  BIC = c(
    BIC(m4_primary),
    BIC(m4_comp),
    BIC(m4_explor),
    BIC(m4_total)
  ),
  McFadden = c(
    pR2(m4_primary)["McFadden"],
    pR2(m4_comp)["McFadden"],
    pR2(m4_explor)["McFadden"],
    pR2(m4_total)["McFadden"]
  )
)

model_compare %>%
  mutate(across(where(is.numeric), ~ round(., 3))) %>%
  kable()
```

## Predicted probabilities for the primary interaction model

```{r predicted-probabilities-primary}
newdat_primary <- expand.grid(
  z_neigh_disadv = 0,
  gender = levels(dat_binary$gender)[1],
  income = levels(dat_binary$income)[1],
  z_age = c(-1, 0, 1),
  z_reward = c(-1, 0, 1),
  z_depression = 0,
  z_mania = 0,
  z_anxiety = 0,
  z_loneliness = c(-1, 0, 1)
)

# Add DOSPERT terms at the mean for prediction.
for (term in primary_dospert_terms) {
  newdat_primary[[term]] <- 0
}

newdat_primary$pred_prob <- predict(m4_primary, newdata = newdat_primary, type = "response")
newdat_primary
```

```{r interaction-plot-primary}
newdat_primary %>%
  filter(z_age == 0) %>%
  mutate(
    Reward = factor(z_reward, labels = c("Low (-1 SD)", "Mean", "High (+1 SD)")),
    Loneliness = factor(z_loneliness, labels = c("Low (-1 SD)", "Mean", "High (+1 SD)"))
  ) %>%
  ggplot(aes(x = Reward, y = pred_prob, group = Loneliness, color = Loneliness)) +
  geom_point(size = 3) +
  geom_line() +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(
    x = "Reward sensitivity",
    y = "Predicted P(PG)",
    color = "Loneliness",
    title = "Primary model: Reward sensitivity × loneliness"
  ) +
  theme_minimal(base_size = 13)
```

## Secondary analysis: severity within PG only

```{r build-pg-severity-dataset}
dat_pg <- df %>%
  filter(pg_group == "PG") %>%
  drop_na(
    gender, income_median_split,
    TotalCrime, unemployed, Gini,
    demo_yrs,
    rew_composite,
    dospert_gam, dospert_inv, dospert_hs, dospert_soc, dospert_ethic, dospert_rec,
    susd_depression, susd_mania, scared_total, uclal_total,
    gsas_total
  ) %>%
  mutate(
    income = income_median_split,
    z_crime = z_score(TotalCrime),
    z_unemp = z_score(unemployed),
    z_gini  = z_score(Gini),
    z_neigh_disadv = rowMeans(cbind(z_crime, z_unemp, z_gini), na.rm = TRUE),
    z_age = z_score(demo_yrs),
    z_reward = z_score(rew_composite),
    z_dospert_gam   = z_score(dospert_gam),
    z_dospert_inv   = z_score(dospert_inv),
    z_dospert_hs    = z_score(dospert_hs),
    z_dospert_soc   = z_score(dospert_soc),
    z_dospert_ethic = z_score(dospert_ethic),
    z_dospert_rec   = z_score(dospert_rec),
    dospert_nongam_primary_raw = if (include_investment_primary) {
      rowMeans(cbind(dospert_hs, dospert_soc, dospert_ethic, dospert_rec, dospert_inv), na.rm = FALSE)
    } else {
      rowMeans(cbind(dospert_hs, dospert_soc, dospert_ethic, dospert_rec), na.rm = FALSE)
    },
    z_dospert_nongam_comp = z_score(dospert_nongam_primary_raw),
    z_depression = z_score(susd_depression),
    z_mania      = z_score(susd_mania),
    z_anxiety    = z_score(scared_total),
    z_loneliness = z_score(uclal_total),
    z_gsas_total = z_score(gsas_total)
  ) %>%
  filter(gender %in% c("Male", "Female"))

cat("PG severity analytic N =", nrow(dat_pg), "\n")
```

```{r severity-block-models}
# Severity models mirror the logistic structure, but the outcome is GSAS total.
severity_formula_1 <- make_formula("gsas_total", list(block_1))
severity_formula_2 <- make_formula("gsas_total", list(block_1, block_2_primary))
severity_formula_3 <- make_formula("gsas_total", list(block_1, block_2_primary, block_3))
severity_formula_4 <- make_formula("gsas_total", list(block_1, block_2_primary, block_3), interactions = block_4_interactions)

s1 <- lm(severity_formula_1, data = dat_pg)
s2 <- lm(severity_formula_2, data = dat_pg)
s3 <- lm(severity_formula_3, data = dat_pg)
s4 <- lm(severity_formula_4, data = dat_pg)
```

```{r severity-model-comparison}
anova(s1, s2, s3, s4)
AIC(s1, s2, s3, s4)
BIC(s1, s2, s3, s4)
summary(s4)
```

```{r severity-vif}
vif(s3)
vif(s4, type = "predictor")
```

```{r severity-clean-table}
tidy(s4, conf.int = TRUE) %>%
  mutate(
    across(c(estimate, std.error, conf.low, conf.high), ~ round(., 3)),
    p.value = round(p.value, 3),
    sig = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      p.value < .10  ~ ".",
      TRUE ~ ""
    )
  ) %>%
  kable()
```

## Notes for interpretation

```{r notes, echo=FALSE, results='asis'}
cat(
"\nSuggested read of this notebook:\n",
"\n1. Start with the primary logistic model using non-gambling DOSPERT domains.\n",
"\n2. Check whether Block 4 materially improves fit. If not, the main-effects model may be the cleaner result to report.\n",
"\n3. Compare the primary model with the non-gambling composite and exploratory gambling-domain models.\n",
"\n4. Treat the severity models as a different question: what predicts symptom burden among PG cases, not what distinguishes cases from controls.\n",
"\n5. If investment/financial risk looks too gambling-adjacent, keep include_investment_primary = FALSE.\n"
)
```
